---
title: "MultLR"
editor: visual
---

## Example

Looking at the `penguins` from `palmerpenguins`, fit a model between `body_mass_g` as an outcome variable and `flipper_length_mm` and `bill_length_mm` as predictor variables.

```{r}
library(palmerpenguins)
penguins |>  lm(body_mass_g ~ flipper_length_mm + bill_length_mm, 
                data = _)
```

```{r}
unique(penguins$island)
```

## Fitting a model with categorical variables

Looking at the `penguins` from `palmerpenguins`, fit a model between `body_mass_g` as an outcome variable and `flipper_length_mm` and `islands` as predictor variables.

```{r}
library(palmerpenguins)
penguins |>  lm(body_mass_g ~ flipper_length_mm + island,
                data = _)
```

```{r}
#penguins <- penguins |> 
 # mutate(island_dream = relevel(island , ref = "Dream"))
#penguins |> lm(body_mass_g ~ flipper_length_mm + island_dream, 
#               data = _)
```

```{r}
#library(tidyverse)
#penguins|>ggplot(aes(body_mass_g,island))+geom_boxplot()
```

```{r}
x1 <- rnorm(1000, 2, 1)
x2<- rnorm(1000,-4,1)
error<-rnorm(1000,0,2)

y<-rep(0,length(x1))
y<-3+2*x1+4*x2+error


```

## Fit Model

Fit a model between $Y$ and $X_1$.

Repeat the process 1000 times. and answer the following questions:

-   On average does $\beta_1$ get estimated correctly? Why?

-   What is the average model variance?

```{r}
beta_1<-rep(0,1000)
sd<-rep(0,1000)
for (i in 1:1000){
  x1 <- rnorm(1000, 2, 1)
x2<- rnorm(1000,-4,1)
error<-rnorm(1000,0,2)

y<-rep(0,length(x1))
y<-3+2*x1+4*x2+error
slm<-lm(y~x1)
beta_1[i]<-coef(slm)[2]
sd[i]<-sigma(slm)}

mean(beta_1)
mean(sd)
```

## MLR Model

Instead of fitting a simple linear regression model. Fit a model that will include predictor $X_2$. This can be done by adding $X_2$ in R:

```{r}
beta<-matrix(rep(0,3000),1000,3)
dim(beta)
  sd_mlr<-rep(0,1000)
for(i in 1:1000){
  x1 <- rnorm(1000, 2, 1)
x2<- rnorm(1000,-4,1)
error<-rnorm(1000,0,2)

y<-rep(0,length(x1))
y<-3+2*x1+4*x2+error
mlr<-lm(y ~ x1 + x2)
beta[i,] <- coef(mlr)
#beta
sd_mlr[i]<-sigma(mlr)
#sd_mlr
}
  colMeans(beta)
  mean(sd_mlr)

```

Modify your simulation study and see what happens to $\beta_1$ and the model variance.
